Constraining the Interiors of Asteroids Through Close Encounters

### 2022-03-01 summaries/21-11/21-11.tex
Lumpy model written

\subsection{Harmonic model}
We now explore a model that seeks to narrow down the space of allowed density distributions a different way than the likelihood model; we define the ``nice'' density distributions directly. They are the density distributions with zero Laplacian: $\nabla^2 \rho = 0$ inside the asteroid. The expansion of such a function in spherical coordinates is given by 
\begin{equation}
  \rho(\bm r) = \sum_{\ell, m} D_{\ell m} R_{\ell m}^*(\bm r)
  \label{eqn:harmonic-rho}
\end{equation}
where the terms which lead to $\rho \rightarrow \infty$ at the origin have been removed and $D_{\ell m}$ are free parameters. The complex conjugate was added for convenience. By setting a maximum on $\ell$, we restrict the number of degrees of freedom to $(\ell_\mathrm{max}+ 1)^2$. Choosing the same maximum $\ell$ as the maximum $\ell$ for the $H_{\ell m}$ moments, we have $m=n-1$ coefficients $D_{\ell m}$ which can be stacked into an $m$-dimensional vector $\bm D$.

Inserting equation \ref{eqn:harmonic-rho} into equations \ref{eqn:hlm-constraint} and \ref{eqn:b-constraint}, we get 
\begin{equation}
  K_{\ell', m'} = \sum_{\ell m} \frac{1}{\mu_m a_m^{\ell'}} D_{\ell m} \int_\mathcal{A} d^3 r R_{\ell m}^*(\bm r) R_{\ell' m'}(\bm r)
\end{equation}
This is an over-determined matrix equation $\bm K = M \bm D$, where $M$ is an $n \times m$ matrix. The Moore-Penrose inverse can therefore be used again to find an inverse $A=M^+$ which yields approximately correct $\bm D$ (approximate in that the norm of the error vector between $\bm K$ and $M \bm D$ is minimized). However, the form of the inverse changes due to the matrix now being taller than it is wide. The Moore-Penrose is now
\begin{equation}
  M^+ = (M^*M)^{-1} M^*
  \label{eqn:mpi-overdetermined}
\end{equation}
and the model concludes with $A=M^+$.

@@

The prime in equation \ref{eqn:klm-stack} denotes that the moments are calculated in the hybrid frame. We can relate the primed moments to the moments calculated relative to each body's center of mass via the translation rules for solid spherical harmonics:
\begin{equation}
  K_{\ell m}'^i = \sum_{\ell' m'} (-1)^{\ell - \ell'} R_{\ell - \ell', m - m'}(\bm x_i) K_{\ell' m'}^i \qquad a_i'^2 = a_i^2 + x_i^2 K_{00}
  \label{eqn:translate-klm}
\end{equation}
Here, $K_{00}^i$ is the added mass $\mu_i$ of lump $i$, while $K_{1m}=0$ and $K_{2m}$ incorporate its orientation and moment of inertia ratios. The volume is constrained by $a_i^2$. These values map directly onto an ellipsoid representation via equations \ref{eqn:ellipsoid-klm}, so that if $K_{\ell m}^i$, $a_i^2$, $\mu_i$, and $x_i$ are known, then the density distribution is known. Note that if $\bm x_i$ is known, then equation \ref{eqn:translate-klm} are linear in $K_{\ell m}^i$. Therefore, if we define $\bm D$ to contain $K_{2m}^i$ and $a_i^2$ in some order, then equations \ref{eqn:translate-klm} and \ref{eqn:klm-stack} define a matrix equation $\bm K = \bm C + M \bm D$ which can be solved by setting $A$ equal to the Moore-Penrose inverse of $M$. Henceforth, we assume that the dimensions of $\bm D$ are less than or equal to the number of known parameters (the shaded regions of table \ref{tab:lump-dof}), so that $A$ is given by equation \ref{eqn:mpi-overdetermined}). Note that the $\bm C$ term of this equation came from the expansion of $K_{\ell m}'^0$ and $a_0^2$, which is written in terms of the already-known displacement of the asteroid surface from its center of mass.

This model's strategy will be to choose $\bm x_i$ and $\mu_i$ via a nonlinear process, then fit for the values $K_{\ell m}^i$ and $a_i^2$ using the linear format described in the introduction to this section. The constraints on $\bm x_i$ and $|mu_i$ are primarily
\begin{equation}
  \sum_{i=1}^N \bm x_i \mu_i = \mu_0 \bm x_0
  \label{eqn:lump-constraints}
\end{equation}
where $\bm x_0$ is the known displacement of the centroid of the asteroid model from the center of mass. There are also additional constraints, such as that $\bm x_i$ should lie inside the asteroid which can be enforced manually. The overall mass of the asteroid is $\mu_m = \mu_0 + \sum_{i=1}^N \mu_i$, and has been set, so we may eliminate $\mu_0$ from equation \ref{eqn:lump-constraints}. The constraint is now a function of $\mu_i$ and $\bm x_i$ for $i \geq 1$. The matrix equation defining $M$ is overdetermined, but we would like it to have a solution nevertheless. We therefore solve for $\bm x_i$ and $\mu_i$ to yield an $M$ with solutions. This is done by minimizing
\begin{equation}
  |(M(\bm x_i, \mu_i) M^+(\bm x_i, \mu_i) - \mathds{1}) (\bm K - \bm C)|^2.\
  \label{eqn:lump-minimize}
\end{equation}
If $\bm x_i$ and $\mu_i$ are chosen to minimize this function, with the constraint of \ref{eqn:lump-constraints} obeyed, then $A=M^+$ and $\bm B(\bm r)$ following from the resulting values of $K_{\ell m}^i$ and $a_i^2$ defines the model. If the resulting density distribution is somehow excluded (it predicts negative density distributions, the lumps extend outside the asteroid, etc.), then another minimum of equation \ref{eqn:lump-minimize} can be used, or another combination of $N$ and $d$ listed in table \ref{tab:lump-dof}.

>>>>

\subsection{Harmonic model}
We now explore a model that seeks to narrow down the space of allowed density distributions a different way than the likelihood model; we define the ``nice'' density distributions directly. They are the density distributions with zero Laplacian: $\nabla^2 \rho = 0$ inside the asteroid. The expansion of such a function in spherical coordinates is given by 
\begin{equation}
  \rho(\bm r) = \sum_{\ell, m} \frac{D_{\ell m}}{a_m^\ell} R_{\ell m}^*(\bm r)
  \label{eqn:harmonic-rho}
\end{equation}
where the terms which lead to $\rho \rightarrow \infty$ at the origin have been removed and $D_{\ell m}$ are free parameters. The complex conjugate was added for convenience. By setting a maximum on $\ell$, we restrict the number of degrees of freedom to $(\ell_\mathrm{max}+ 1)^2$. Choosing the same maximum $\ell$ as the maximum $\ell$ for the $H_{\ell m}$ moments, we have $m=n-1$ coefficients $D_{\ell m}$ which can be stacked into an $m$-dimensional vector $\bm D$.

Inserting equation \ref{eqn:harmonic-rho} into equations \ref{eqn:hlm-constraint} and \ref{eqn:b-constraint}, we get 
\begin{equation}
  K_{\ell', m'} = \sum_{\ell m} \frac{1}{\mu_m a_m^{\ell'} a_m^\ell} D_{\ell m} \int_\mathcal{A} d^3 r R_{\ell m}^*(\bm r) R_{\ell' m'}(\bm r)
  \label{eqn:harmonic-mat}
\end{equation}
This is an over-determined matrix equation $\bm K = M \bm D$, where $M$ is an $n \times m$ matrix. The Moore-Penrose inverse can therefore be used again to find an inverse $A=M^+$ which yields approximately correct $\bm D$ (approximate in that the norm of the error vector between $\bm K$ and $M \bm D$ is minimized). However, the form of the inverse changes due to the matrix now being taller than it is wide. The Moore-Penrose is now
\begin{equation}
  M^+ = (M^*M)^{-1} M^*
  \label{eqn:mpi-overdetermined}
\end{equation}
and the model concludes with $A=M^+$.

Note that, in the special case where the asteroid is a sphere of radius $R$, the matrix defined by equation \ref{eqn:harmonic-mat} is diagonal, with entries 
\begin{equation}
  M_{\ell m; \ell' m'} = \frac{4\pi R^3}{\mu_m} \frac{R^{2\ell}}{a_m^{2\ell}} \frac{\delta_{\ell \ell'} \delta_{m m'}}{(4\ell^2 + 8\ell + 3)(\ell - m)!(\ell+m)!}.
\end{equation}
A nonspherical perturbation will introduce small off-diagonal entries to $M$. This diagonal represents an interpretation of the physical meaning of $K_{\ell m}$; they are directly related through this form of $M$ and the matrix equation $\bm K = M \bm D$ to the coefficients of the spherical harmonic expansion of the asteroid in the spherical case.

@@

The prime in equation \ref{eqn:klm-stack} denotes that the moments are calculated in the hybrid frame. We can relate the primed moments to the moments calculated relative to each body's center of mass via the translation rules for solid spherical harmonics:
\begin{equation}
  K_{\ell m}'^i = \sum_{\ell' m'} (-1)^{\ell - \ell'} R_{\ell - \ell', m - m'}(\bm x_i) K_{\ell' m'}^i \qquad a_i'^2 = a_i^2 + x_i^2 K_{00}
  \label{eqn:translate-klm}
\end{equation}
The dummy indices $\ell', m'$ should only be summed over values in which $\ell-\ell' \geq 0$ and $|m-m'| \leq \ell - \ell'$. Here, $K_{00}^i$ is the added mass $\mu_i$ of lump $i$, while $K_{1m}=0$ and $K_{2m}$ incorporate its orientation and moment of inertia ratios. The volume is constrained by $a_i^2$. These values map directly onto an ellipsoid representation via equations \ref{eqn:ellipsoid-klm}, so that if $K_{\ell m}^i$, $a_i^2$, $\mu_i$, and $x_i$ are known, then the density distribution is known. Note that if $\bm x_i$ is known, then equation \ref{eqn:translate-klm} are linear in $K_{\ell m}^i$. Therefore, if we define $\bm D$ to contain $K_{2m}^i$ and $a_i^2$ in some order, then equations \ref{eqn:translate-klm} and \ref{eqn:klm-stack} define a matrix equation $\bm K = \bm C + M \bm D$ which can be solved by setting $A$ equal to the Moore-Penrose inverse of $M$. Henceforth, we assume that the dimensions of $\bm D$ are less than or equal to the number of known parameters (the shaded regions of table \ref{tab:lump-dof}), so that $A$ is given by equation \ref{eqn:mpi-overdetermined}). Note that the $\bm C$ term of this equation came from the expansion of $K_{\ell m}'^0$ and $a_0^2$, which is written in terms of the already-known displacement of the asteroid surface from its center of mass.

This model's strategy will be to choose $\bm x_i$ and $\mu_i$ via a nonlinear process, then fit for the values $K_{\ell m}^i$ and $a_i^2$ using the linear format described in the introduction to this section. The constraints on $\bm x_i$ and $|mu_i$ are primarily
\begin{equation}
  \sum_{i=1}^N \bm x_i \mu_i + \mu_0 \bm x_0 = 0
  \label{eqn:lump-constraints}
\end{equation}
where $\bm x_0$ is the known displacement of the centroid of the asteroid model from the center of mass. There are also additional constraints, such as that $\bm x_i$ should lie inside the asteroid which can be enforced manually. The overall mass of the asteroid is $\mu_m = \mu_0 + \sum_{i=1}^N \mu_i$, and has been set, so we may eliminate $\mu_0$ from equation \ref{eqn:lump-constraints}. The constraint is now a function of $\mu_i$ and $\bm x_i$ for $i \geq 1$. The matrix equation defining $M$ is overdetermined, but we would like it to have a solution nevertheless. We therefore solve for $\bm x_i$ and $\mu_i$ to yield an $M$ with solutions. This is done by minimizing
\begin{equation}
  |(M(\bm x_i, \mu_i) M^+(\bm x_i, \mu_i) - \mathds{1}) (\bm K - \bm C)|^2.\
  \label{eqn:lump-minimize}
\end{equation}
If $\bm x_i$ and $\mu_i$ are chosen to minimize this function, with the constraint of \ref{eqn:lump-constraints} obeyed, then $A=M^+$ and $\bm B(\bm r)$ following from the resulting values of $K_{\ell m}^i$ and $a_i^2$ defines the model. If the resulting density distribution is somehow excluded (it predicts negative density distributions, the lumps extend outside the asteroid, etc.), then another minimum of equation \ref{eqn:lump-minimize} can be used, or another combination of $N$ and $d$ listed in table \ref{tab:lump-dof}.
